---
title: "SEER Medicaid"
author: "Kim Johnson"
date: '2022-06-06'
output: html_document
---
# What this analysis will do from project description
1.	Is there an association between cancer stage at diagnosis and survival and insurance status at diagnosis in children (0 to 14 years) and adolescents and young adults (15-39 years) overall and stratified by five-year age groups?

Hypothesis: We hypothesize that those with Medicaid will have an increased risk of death and later-stage diagnosis compared to those with private insurance and that this association will become progressively stronger with increasing age. 

2.	Determine whether sociodemographic and economic factors modify associations.

Hypothesis: We hypothesize that certain sociodemographic and economic factors (e.g. race/ethnicity, rural/urban status, census tract SES measures) modify associations between insurance and stage at diagnosis and survival. 

3.	Determine whether associations between stage at diagnosis and survival differ by the timing of Medicaid enrollment with the following categories: 
o	before diagnosis defined as those who enrolled before the year of diagnosis, 
o	around diagnosis defined as in the year of diagnosis 
o	after diagnosis defined as in the year after diagnosis
o	Not enrolled (did not link to Medicaid)*

Hypothesis: We hypothesize that those enrolled in Medicaid around diagnosis or after diagnosis will have later stage diagnoses and a greater hazard of death than those enrolled before diagnosis.

# Further notes on timing of enrollment
1. They considered them enrolled in Medicaid within a 25 month window so not just the flag for Medicaid (I think there was refinement after that to limit it to 25 months)
2. Those that were enrolled in Medicaid in more than one state during that 25 month window were excluded and also those with a match score of <5.
3. Then in that group, it appears that they probably made the following variable or variables:
Enrolled any month before the diagnosis month but not during the dx month==enrolled before dx
Enrolled in the month of dx  == enrolled at dx
 
Those enrolled before and after were classified as enrolled before dx.
 
So this does not get at churning (yet) but I think may be more straightforward than what I was thinking. I would hypothesize that the first bulleted group has better survival than the second.
# Open libraries
```{r}
pacman::p_load(readr, dplyr, tidyverse, lubridate, survival, survminer, hablar)
```

# Import files sent by SEER Medicaid and get data into right columns, assign column names
```{r, eval =FALSE}
# # Medicaid enrollment--imports fixed width columns and gives them names
# enroll <- read_fwf('medicaid.enroll.txt', 
#                fwf_cols(patient_id = 15, STATE_CD = 2, YEAR = 4, DOB_MATCH_SCORE_HI = 1, SEX_MATCH_SCORE_HI = 1,
#                        ID_MATCH_SCORE_HI=1, MATCH_SCORE_HI=2, MEDICAID_ELIG_CD1=1, MEDICAID_ELIG_CD2=1,
#                       MEDICAID_ELIG_CD3=1, MEDICAID_ELIG_CD4=1, MEDICAID_ELIG_CD5=1, MEDICAID_ELIG_CD6=1,
#                       MEDICAID_ELIG_CD7=1, MEDICAID_ELIG_CD8=1, MEDICAID_ELIG_CD9=1, MEDICAID_ELIG_CD10=1,
#                         MEDICAID_ELIG_CD11=1, MEDICAID_ELIG_CD12=1), 
#                skip = 0)

# # SEER file 
# 
# seer<-read_fwf('medicaid.cancer 2.txt', na=".", fwf_positions(
# c(1, 16,18,20,21,22,24,25,28,31,33,35,39,41,45,46,50,51,55,56,57,58,59,62,64,66,67,69, 71, 72,73,74,75,76,77,78,79,80,81,82,83,84,86,88,90,92,94,95,96,97,98,99,100,103,106,109, 111,114,117,120,123,126,129,132,135,138,141,144,147,150,153,156, 159,161,163,165,167,168,174,180,186,188,189,190,192,193,194,195,196,197,199, 200,201,203,205,210,213,216,217,219,224,227,228,229,230,231,233,235,236,238,241,242, 243,248, 253, 254, 255,256, 258,263,264,265,266,267,268,269,270,271,273,279,282,285,288,291,293,295,297,299, 300,301,302,303,304,305,306,308,310,316,347,353,359,365,366,367,368,369,378,380,382,384,388,401,402,403,404,405, 406,408,412,414, 419, 423,425,427,429, 431,433,437,438,439,440,441,442,443,445, 447, 448, 452, 456, 460, 464,465,466,467,468,469,470, 471,472,473,474,475,476,477, 479, 480,485,490,495,500,505, 510,511,512,513,514,515,519,520, 521,524,525,526,530,532), 
#   
# c(15,17,19,20,21,23,24,27,30,32,34,38,40,44,45,49,50,54,55,56,57,58,61,63,65,66,68,70,71, 72,73,74,75,76,77,78,79,80,81,82,83,84,86,88,90,93,94,95,96,97,98, 99,102,105,108,110,113,116,119,122,125,128,131,134,137, 140, 143,146,149,152,155,158,160,162,164,166,167,173,179, 185,187,188,189,191, 192,193,194,195,195, 198,199, 200,202,204,209,212,215,216,218,220, 226,227,228,229,230,232,234,235,237,240,241,242,247, 252, 253, 254,255,257, 259, 263,264,265,266,267,268,269,270, 272,273,281,284,287,290,292,294,295,298,299, 300, 301,302,303,304,305,307, 309,311,316,352,358, 364,365,366,367,368,369, 379,381,383,387,391,401,402,403,404,405, 407,411,413,417, 422,424,426,428,430, 432,436,437,438,439,440,441,442,444,446, 447, 451, 455,459,463,464, 465, 466,467,468,469,470, 471,472,473,474,475,476,477,479, 484,489,494,499,504,509,510,511,512,513,514,515,519,520, 523,524,525,529,531,534)))
# 
# colnames(seer)<- c("patient_id", "SEER_registry", "SEERregistrywithCAandGAaswholes", 
# "Louisiana20051stvs2ndhalfofyear", "Marital_status_at_diagnosis", 
# "Race_ethnicity", "sex", "Agerecodewithsingleages_and_100", "agerecodewithsingle_ages_and_85", 
# "Sequence_number", "Month_of_diagnosis", "Year_of_diagnosis", 
# "Month_of_diagnosis_recode", "Primary_Site", "Laterality", "Histology_ICD_O_2", 
# "Behavior_code_ICD_O_2", "Histologic_Type_ICD_O_3", "Behavior_code_ICD_O_3", 
# "Grade", "Diagnostic_Confirmation", "Type_of_Reporting_Source", 
# "EOD_10_size_1988_2003", "EOD_10_extent_1988_2003", "EOD10Prostatepath_ext_1995_2003", 
# "EOD_10_nodes_1988_2003", "Regional_nodes_positive_1988", "Regional_nodes_examined_1988", 
# "Expanded_EOD_1_CP53_1973_1982", "Expanded_EOD_2_CP54_1973_1982", 
# "Expanded_EOD_3_CP55_1973_1982", "Expanded_EOD_4_CP56_1973_1982", 
# "Expanded_EOD_5_CP57_1973_1982", "Expanded_EOD_6_CP58_1973_1982", 
# "Expanded_EOD_7_CP59_1973_1982", "Expanded_EOD_8_CP60_1973_1982",
# "Expanded_EOD_9_CP61_1973_1982", "Expanded_EOD_10_CP62_1973_1982",
# "Expanded_EOD_11_CP63_1973_1982", "Expanded_EOD_12_CP64_1973_1982",
# "Expanded_EOD_13_CP65_1973_1982", "V2Digit_NS_EOD_part_1_1973_1982",
# "V2Digit_NS_EOD_part_2_1973_1982", "V2Digit_SS_EOD_part_1_1973_1982",
# "V2Digit_SS_EOD_part_2_1973_1982", "EOD_4_size_1983_1987", "EOD_4_extent_1983_1987",
# "EOD_4_nodes_1983_1987", "Coding_system_EOD_1973_2003", "Tumor_marker_1_1990_2003",
# "Tumor_marker_2_1990_2003", "Tumor_marker_3_1998_2003", "CS_tumor_size_2004_2015",
# "CS_extension_2004_2015", "CS_lymph_nodes_2004_2015", "CS_mets_at_dx_2004_2015",
# "CSsitespecificfactor12004varyin", "CSsitespecificfactor22004varyin",
# "CSsitespecificfactor32004varyin", "CSsitespecificfactor42004varyin",
# "CSsitespecificfactor52004varyin", "CSsitespecificfactor62004varyin",
# "CSsitespecificfactor72004varyin", "CSsitespecificfactor82004varyin",
# "CSsitespecificfactor92004varyin", "CSsitespecificfactor102004varyi",
# "CSsitespecificfactor112004varyi", "CSsitespecificfactor122004varyi",
# "CSsitespecificfactor132004varyi", "CSsitespecificfactor152004varyi",
# "CSsitespecificfactor162004varyi", "CSsitespecificfactor252004varyi",
# "Derived_AJCC_T_6th_ed_2004_2015", "Derived_AJCC_N_6th_ed_2004_2015",
# "Derived_AJCC_M_6th_ed_2004_2015", "DerivedAJCCStageGroup6thed20042",
# "SEERCombinedSummaryStage2000200", "CSversioninputoriginal2004_2015",
# "CS_version_derived_2004_2015", "CSversioninputcurrent_2004_2015",
# "RX_Summ_Surg_Prim_Site_1998", "RX_Summ_Scope_Reg_LN_Sur_2003",
# "RX_Summ_Surg_Oth_Reg_Dis_2003", "RXSummReg_LN_Examined_1998_2002",
# "RX_Summ_Systemic_Surg_Seq", "RX_Summ_Surg_Rad_Seq", "Reasonnocancer_directed_surgery",
# "Radiation_recode", "Chemotherapy_recode_yes_no_unk", "Sitespecificsurgery19731997vary",
# "Scopeofreglymphndsurg_1998_2002", "Surgeryofothregdissites19982002",
# "Record_number_recode", "Age_recode_with_1_year_olds", "Site_recode_ICD_O_3_WHO_2008",
# "ICCCsiterecode_ICD_O_3_WHO_2008", "ICCCsiterecextendedICDO3WHO2008",
# "Behavior_recode_for_analysis", "Histologyrecode_broad_groupings",
# "Histologyrecode_Brain_groupings", "TNM_7_CS_v0204_Schema", "Race_recode_White_Black_Other",
# "Race_recode_W_B_AI_API", "OriginrecodeNHIAHispanicNonHisp",
# "SEER_historic_stage_A_1973_2015", "AJCCstage_3rd_edition_1988_2003",
# "SEERmodifiedAJCCstage3rd1988200", "Firstmalignantprimary_indicator",
# "state", "county", "PRCDA_2017", "PRCDA_Region", "COD_to_site_recode",
# "COD_to_site_rec_KM", "Vitalstatusrecodestudycutoffuse", "IHS_Link",
# "Summary_stage_2000_1998", "AYA_site_recode_WHO_2008", "Lymphomasubtype_recode_WHO_2008",
# "SEERcausespecificdeathclassific", "SEERothercauseofdeathclassifica",
# "CSTumor_Size_Ext_Eval_2004_2015", "CS_Reg_Node_Eval_2004_2015",
# "CS_Mets_Eval_2004_2015", "Primary_by_international_rules", "ERStatusRecodeBreastCancer_1990",
# "PRStatusRecodeBreastCancer_1990", "CS_Schema_AJCC_6th_Edition",
# "LymphvascularInvasion2004varyin", "Derived_AJCC_T_7th_ed_2010",
# "Derived_AJCC_N_7th_ed_2010", "Derived_AJCC_M_7th_ed_2010", "DerivedAJCCStageGroup7thed_2010",
# "BreastAdjustedAJCC6thT1988_2015", "BreastAdjustedAJCC6thN1988_2015",
# "BreastAdjustedAJCC6thM1988_2015", "BreastAdjustedAJCC6thStage19882",
# "Derived_HER2_Recode_2010", "Breast_Subtype_2010", "LymphomaAnnArborStage_1983_2015",
# "SEERCombinedMetsat_DX_bone_2010", "SEERCombinedMetsatDX_brain_2010",
# "SEERCombinedMetsatDX_liver_2010", "SEERCombinedMetsat_DX_lung_2010",
# "TvaluebasedonAJCC_3rd_1988_2003", "NvaluebasedonAJCC_3rd_1988_2003",
# "MvaluebasedonAJCC_3rd_1988_2003", "RadiationtoBrainorCNSRecode1988",
# "Census_Tract_1990", "Census_Tract_2000", "Census_Tract_2010",
# "Census_Coding_System", "Census_Tract_Certainty_1990", "Census_Tract_Certainty_2000",
# "Census_Tract_Certainty_2010", "Census_Tract_Poverty_Indicator",
# "Rural_Urban_Continuum_Code_1993", "Rural_Urban_Continuum_Code_2003",
# "Rural_Urban_Continuum_Code_2013", "Health_Service_Area", "HealthService_Area_NCI_Modified",
# "Derived_SS1977_flag", "Derived_SS2000_flag", "Firstcourseofreconstruct1998200",
# "Radiation", "RadiationtoBrainorCNS_1988_1997", "SEER_DateofDeath_Month",
# "SEER_DateofDeath_Year", "Month_of_last_follow_up_recode", "Year_of_last_follow_up_recode",
# "Year_of_birth", "Date_of_diagnosis_flag", "Date_therapy_started_flag",
# "Date_of_birth_flag", "Date_of_last_follow_up_flag", "Month_therapy_started",
# "Year_therapy_started", "Other_cancer_directed_therapy", "Derived_AJCC_flag",
# "Derived_SS1977", "Derived_SS2000", "SEER_Summary_stage_1977_9500",
# "SEER_Summary_stage_2000_0103", "NAPIIA_Derived_API_Race", "Primary_Payer_at_DX",
# "COD_ICD_code", "Cause_of_Death_ICD_8_or_9", "Cause_of_Death_ICD_10",
# "Recode_ICD_0_2_to_9", "Recode_ICD_0_2_to_10", "NHIA_Derived_Hisp_Origin",
# "Age_site_edit_override", "Sequencenumber_dx_conf_override",
# "Site_type_lat_seq_override", "Surgerydiagnostic_conf_override",
# "Site_type_edit_override", "Histology_edit_override", "Report_source_sequence_override",
# "Seq_ill_defined_site_override", "LeukLymphdxconfirmationoverride",
# "Site_behavior_override", "Site_EOD_dx_date_override", "Site_laterality_EOD_override",
# "Site_laterality_morph_override", "Insurance_Recode_2007", "Yost_ACS_2006_2010",
# "Yost_ACS_2010_2014", "Yost_ACS_2013_2017", "Yost_ACS_2006_2010_State_based",
# "Yost_ACS_2010_2014_State_based", "Yost_ACS_2013_2017_State_based",
# "Yost_ACS_2006_2010_quintile", "Yost_ACS_2010_2014_quintile",
# "Yost_ACS_2013_2017_quintile", "YostACS20062010quintileStatebas",
# "YostACS20102014quintileStatebas", "YostACS20132017quintileStatebas",
# "HPV_recode_2010", "In_Medicaid_Flag", "OncotypeDXBreastRecurrenceScore",
# "OncotypeDXRSgroupRS18RS1830RS30", "OncotypeDXreasonno_score_linked",
# "Oncotype_DX_year_of_test_report", "OncotypeDX_month_of_test_report",
# "OncotypeDXmonthssince_diagnosis")

# hit control shift c while highlighting above to uncomment/comment section
```

# Create dataframes for Medicaid enrollment and SEER for export that are limited to those diagnosed at 0 to 39 years
# Exclude those over 39 years old and export data to make the datasets more manageeable on import
```{r}
# exclude those 40 and over from seer first
# seer_0to39<-seer %>%
#   filter(Agerecodewithsingleages_and_100<40)

# get ids from seer file to limit enrollment file

# ids<-seer_0to39$patient_id

# enroll_seerids<-enroll %>% # limit Medicaid enrollment file to SEER ids
 # filter(patient_id %in% ids)

```

# Export these data as csv files for quicker import
```{r}
# write.csv(seer_0to39,"seer0to39.csv", row.names = FALSE)
 
# write.csv(enroll_seerids,"enroll_seerids.csv", row.names = FALSE)
```

# Import files created with text commented out above for analysis
```{r}
seer0to39<-read_csv("/Volumes/kijohnson/Active/seer0to39.csv", show_col_types = FALSE) # this is 0 to 39 cancer patient data
enroll_seerids<-read_csv("/Volumes/kijohnson/Active/enroll_seerids.csv", show_col_types = FALSE) # this is 0 to 39 enrollee data
mcseer <- read_csv("/Volumes/kijohnson/Active/mcseer_dsb/mcseer_dsb.csv", show_col_types = FALSE) # Derek's file that contains continuity of enrollment variables
```

# Identify ids to exclude where patient has files in multiple states (this is usedd below)
```{r}
df<-enroll_seerids %>% 
   find_duplicates(patient_id, YEAR)
 
x <- unique(df$patient_id) # 841 vector of unique ids to exclude in exclusion section
```

# Reduce canceer dataset to variables needed for analysis
```{r}
# select variables
seer0to39<-seer0to39 %>%
  select("patient_id", "SEER_registry", "Race_ethnicity", "sex", "Agerecodewithsingleages_and_100",
"Sequence_number", "Month_of_diagnosis", "Year_of_diagnosis",
"Month_of_diagnosis_recode", "Primary_Site", "Laterality", "Histology_ICD_O_2",
"Behavior_code_ICD_O_2", "Histologic_Type_ICD_O_3", "Behavior_code_ICD_O_3",
"Grade", "Diagnostic_Confirmation", "Type_of_Reporting_Source", "Derived_AJCC_T_6th_ed_2004_2015", "Derived_AJCC_N_6th_ed_2004_2015","Derived_AJCC_M_6th_ed_2004_2015", "DerivedAJCCStageGroup6thed20042",
"SEERCombinedSummaryStage2000200", "Age_recode_with_1_year_olds", "Site_recode_ICD_O_3_WHO_2008",
"ICCCsiterecode_ICD_O_3_WHO_2008", "ICCCsiterecextendedICDO3WHO2008",
"Behavior_recode_for_analysis", "Histologyrecode_broad_groupings", "Race_recode_White_Black_Other",
"Race_recode_W_B_AI_API", "OriginrecodeNHIAHispanicNonHisp",
"SEER_historic_stage_A_1973_2015", "AJCCstage_3rd_edition_1988_2003",
"SEERmodifiedAJCCstage3rd1988200", "Firstmalignantprimary_indicator",
"state", "county", "PRCDA_2017", "PRCDA_Region", "COD_to_site_recode",
"COD_to_site_rec_KM", "Vitalstatusrecodestudycutoffuse", "IHS_Link",
"Summary_stage_2000_1998", "SEERCombinedSummaryStage2000200", "AYA_site_recode_WHO_2008", 
"SEERcausespecificdeathclassific", "SEERothercauseofdeathclassifica",
"Derived_AJCC_T_7th_ed_2010", "Derived_AJCC_N_7th_ed_2010", "Derived_AJCC_M_7th_ed_2010", "DerivedAJCCStageGroup7thed_2010", "Census_Tract_1990", "Census_Tract_2000", "Census_Tract_2010",
"Census_Coding_System", "Census_Tract_Certainty_1990", "Census_Tract_Certainty_2000",
"Census_Tract_Certainty_2010", "Census_Tract_Poverty_Indicator",
"Rural_Urban_Continuum_Code_1993", "Rural_Urban_Continuum_Code_2003",
"Rural_Urban_Continuum_Code_2013", "Health_Service_Area", "HealthService_Area_NCI_Modified",
"SEER_DateofDeath_Month", "SEER_DateofDeath_Year", "Month_of_last_follow_up_recode", "Year_of_last_follow_up_recode", "Year_of_birth", "Date_of_diagnosis_flag", "Date_therapy_started_flag",
"Date_of_birth_flag", "Date_of_last_follow_up_flag", "Month_therapy_started",
"Year_therapy_started",  "Derived_AJCC_flag","SEER_Summary_stage_1977_9500",
"SEER_Summary_stage_2000_0103", "NAPIIA_Derived_API_Race", "Primary_Payer_at_DX",
"COD_ICD_code", "Cause_of_Death_ICD_8_or_9", "Cause_of_Death_ICD_10",
"Recode_ICD_0_2_to_9", "Recode_ICD_0_2_to_10", "NHIA_Derived_Hisp_Origin",
"Insurance_Recode_2007", "Yost_ACS_2006_2010", "Yost_ACS_2010_2014", "Yost_ACS_2013_2017", "Yost_ACS_2006_2010_State_based", "Yost_ACS_2010_2014_State_based", "Yost_ACS_2013_2017_State_based",
"Yost_ACS_2006_2010_quintile", "Yost_ACS_2010_2014_quintile",
"Yost_ACS_2013_2017_quintile", "YostACS20062010quintileStatebas",
"YostACS20102014quintileStatebas", "YostACS20132017quintileStatebas",
"HPV_recode_2010", "In_Medicaid_Flag")
```

# Data management of variables 
```{r}
# convert primary payer variable to numeric
seer0to39$Primary_Payer_at_DX<-as.numeric(seer0to39$Primary_Payer_at_DX)

# recode to group some categories
seer0to39<-seer0to39%>%
  mutate(insurance =       if_else(In_Medicaid_Flag==0 & Primary_Payer_at_DX %in% c(20,21), 0, # private insurance
                           if_else(In_Medicaid_Flag==1, 1, # ever had Medicaid between 2006 to 2013
                           if_else(In_Medicaid_Flag==0 & Primary_Payer_at_DX %in% c(1,2), 2, # uninsured
                           if_else(In_Medicaid_Flag==0 & Primary_Payer_at_DX %in% c(31,35), 3, # did not link to Medicaid but coded as Medicaid
                           if_else(In_Medicaid_Flag==0 & Primary_Payer_at_DX %in% c(60,61,62,63,64), 4, # Medicare
                           if_else(In_Medicaid_Flag==0 & Primary_Payer_at_DX %in% c(65,66, 67), 5, # Military
                           if_else(In_Medicaid_Flag==0 & Primary_Payer_at_DX == 68, 6, # Indian/Public Health service
                           if_else(In_Medicaid_Flag==0 & Primary_Payer_at_DX == 99, 7, # unknown
                           if_else(In_Medicaid_Flag==0 & Primary_Payer_at_DX == 10, 8, NULL))))))))))  # insurance NOS

# make it a factor variable
seer0to39$insurance<-factor(seer0to39$insurance, levels=c(0:8), labels=c("Private Insurance", "Ever linked to Medicaid", "Uninsured", "Did not link to Medicaid but coded as Medicaid", "Medicare", "Military", "Indian/Public Health Service", "Unknown", "insurance NOS"))

table(seer0to39$insurance, seer0to39$Primary_Payer_at_DX, useNA="always")

# Classification in SEER file
seer0to39<-seer0to39%>%
  mutate(insurance2=if_else(Primary_Payer_at_DX %in% c(20,21), 0, # private insurance
                           if_else(Primary_Payer_at_DX %in% c(1,2), 1, # uninsured
                           if_else(Primary_Payer_at_DX %in% c(31,35), 2, #  Medicaid
                           if_else(Primary_Payer_at_DX %in% c(60,61,62,63,64), 3, # Medicare
                           if_else(Primary_Payer_at_DX %in% c(65,66, 67), 4, # Military
                           if_else(Primary_Payer_at_DX == 68, 5, # Indian/Public Health service
                           if_else(Primary_Payer_at_DX == 99, 6, # unknown
                           if_else(Primary_Payer_at_DX == 10, 7, NULL)))))))))  # insurance NOS

seer0to39$insurance2<-factor(seer0to39$insurance2, levels=c(0:7), labels=c("Private Insurance", "Uninsured", "Medicaid", "Medicare", "Military", "Indian/Public Health Service", "Unknown", "insurance NOS"))

table(seer0to39$Primary_Payer_at_DX,seer0to39$insurance, useNA="always")

# look at insurance at dx by Medicaid flag
table(seer0to39$Primary_Payer_at_DX, seer0to39$In_Medicaid_Flag)
table(seer0to39$Insurance_Recode_2007, seer0to39$In_Medicaid_Flag)
table(seer0to39$Primary_Payer_at_DX, seer0to39$In_Medicaid_Flag)
table(seer0to39$Primary_Payer_at_DX, seer0to39$Insurance_Recode_2007)

# recode demographic variables
# Age, race, hspanic, odeath, cause specfic death
seer0to39<-seer0to39 %>%
  mutate(age_cat=case_when(Age_recode_with_1_year_olds %in% c(0,1) ~ 0, # 0 to 4
                                Age_recode_with_1_year_olds == 2 ~1, # 5 to 9
                                Age_recode_with_1_year_olds == 3 ~2, # 10 to 14
                                Age_recode_with_1_year_olds == 4 ~3, # 15 to 19
                                Age_recode_with_1_year_olds %in% c(5,6,7,8) ~ 4))%>% # 20 to 39
         
         mutate(race_cat=case_when(Race_recode_W_B_AI_API == 1 ~0, # White
                                Race_recode_W_B_AI_API == 2 ~1, # Black
                                Race_recode_W_B_AI_API == 3 ~2, # AIAN
                                Race_recode_W_B_AI_API == 4 ~3, # API
                                Race_recode_W_B_AI_API == 9 ~4)) %>% # Unknown
  
        mutate(hispanic = case_when(OriginrecodeNHIAHispanicNonHisp == 0 ~0, # Non-Hispanic
                                    OriginrecodeNHIAHispanicNonHisp == 1 ~1)) %>% # Hispanic
  
        mutate(race_ethnicity = case_when(race_cat == 0 & hispanic == 0 ~ 0, # Non-Hispanic White
                                race_cat == 1 & hispanic == 0 ~ 1, # Non-Hispanic Black
                                race_cat == 2 & hispanic == 0 ~ 2, # Non-Hispanic AIAN
                                race_cat == 3 & hispanic == 0 ~ 3, # Non-Hispanic API
                                race_cat == 4 & hispanic == 0 ~ 4, # Non-Hispanic Unknown
                                race_cat %in% c(0:4) & hispanic==1 ~5)) %>% # Hispanic
  
        mutate(sex = case_when(sex == 1 ~ 0, # "Male"
                                sex == 2 ~ 1, # "Female"
                                sex == 9 ~ 2)) %>% # "Other"
  
        mutate(any_dead = case_when(Vitalstatusrecodestudycutoffuse == 0 ~ 1, # Dead
                                Vitalstatusrecodestudycutoffuse == 1 ~0)) %>% # Alive
  
        mutate(can_dead = case_when(SEERcausespecificdeathclassific == 0 ~0, # Alive or dead of other cause
                                SEERcausespecificdeathclassific == 1 ~1)) # Dead
# not 8 an 9 are missing or N/A

# Factor variables
seer0to39$age_cat<-factor(seer0to39$age_cat, levels=c(0:4), labels=c("0-4", "5-9", "10-14", "15-19", "20-39"))
seer0to39$race_cat<-factor(seer0to39$race_cat, levels=c(0:4), labels=c("White", "Black", "AIAN", "API", "Unknown"))
seer0to39$hispanic<-factor(seer0to39$hispanic, levels=c(0:1), labels=c("Non-Hispanic", "Hispanic"))
seer0to39$race_ethnicity<-factor(seer0to39$race_ethnicity, levels=c(0:5), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic AIAN", "Non-Hispanic API", "Non-Hispanic Unknown", "Hispanic")) 
seer0to39$any_dead<-factor(seer0to39$any_dead, levels=c(0:1), labels=c("Alive", "Dead"))
seer0to39$sex <- factor(seer0to39$sex, levels = c(0:2), labels = c("Male", "Female", "Not Stated"))

table(seer0to39$age_cat)
table(seer0to39$race_cat)
table(seer0to39$hispanic)
table(seer0to39$race_ethnicity)
table(seer0to39$any_dead)
table(seer0to39$can_dead)
table(seer0to39$sex)

# calculate survival time from month of dx, recode year of dx TO Month_of_last_follow_up_recode, Year_of_last_follow_up_recode. The difference between these variable dates would provide months of follow-up

# SEERothercauseofdeathclassifica
summary(seer0to39$Month_of_diagnosis_recode)
summary(seer0to39$Year_of_diagnosis)
summary(as.numeric(seer0to39$Month_of_last_follow_up_recode))
seer0to39$Month_of_last_follow_up_recode<-as.numeric(seer0to39$Month_of_last_follow_up_recode) # make variable numeric
summary(seer0to39$Month_of_last_follow_up_recode)
summary(seer0to39$Year_of_last_follow_up_recode)

# assign a date of diagnosis using 15 as the day because there is no day of dx so assume mean is 15
seer0to39$proxy_day_of_dx <-15
seer0to39$proxy_day_of_fu <-15

seer0to39<-seer0to39 %>%
  mutate(proxy_dx_date = make_date(Year_of_diagnosis, Month_of_diagnosis_recode, proxy_day_of_dx)) %>% # make date of dx
  mutate(proxy_fu_date = make_date(Year_of_last_follow_up_recode, Month_of_last_follow_up_recode, proxy_day_of_fu)) # make date of follow-up

# calculate the difference in days between the estimated date of dx and estimated date of last follow-up
seer0to39$survday<-as.numeric(difftime(seer0to39$proxy_fu_date, seer0to39$proxy_dx_date, units=c("days")))
summary(seer0to39$survday)

# estimate survival months by dividing by the average number of days in a month
seer0to39$survmo<-seer0to39$survday/30.437

summary(seer0to39$survmo)

# cancer type coding
table(seer0to39$ICCCsiterecode_ICD_O_3_WHO_2008)

class(seer0to39$ICCCsiterecode_ICD_O_3_WHO_2008)

# ICCC3
seer0to39<-seer0to39 %>%
  mutate(can_type = case_when(
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(11, 12,13,14,15)~0,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(21,22,23,24,25)~1,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(31,32,33,34,35,36)~2,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(41,42) ~3,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(50) ~4,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(62,62,63) ~5,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(71,72,75) ~6,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(81,82, 83,84,85) ~7,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(91,92,93,94,95)~8,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(101,102,103, 104,105) ~9,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(111,112,113,114,115,116) ~10,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(121)~11))

seer0to39$can_type<-factor(seer0to39$can_type, levels=c(0:11), labels=c(
  "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases",
  "II. Lymphomas and reticuloendothelial neoplasms",
  "III. CNS and miscellaneous intracranial and intraspinal neoplasms",
  "IV. Neuroblastoma and other peripheral nervous cell tumors",
  "V. Retinoblastoma",
  "VI. Renal tumors",
  "VII. Hepatic tumors",
  "VIII. Malignant bone tumors",
  "IX. Soft tissue and other extraosseous sarcomas",
  "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads",
  "XI. Other malignant epithelial neoplasms and malignant melanomas",
  "XII. Other and unspecified malignant neoplasms"))
    
table(seer0to39$can_type)  

table(seer0to39$AYA_site_recode_WHO_2008)

# create categories for AYA codes
class(seer0to39$AYA_site_recode_WHO_2008) # numeric

seer0to39 <- seer0to39 %>%
  mutate(can_typeAYA = case_when(AYA_site_recode_WHO_2008 %in% c(1,2,3,4) ~ 0,
                                 AYA_site_recode_WHO_2008 %in% c(5,6) ~ 1,
                                 AYA_site_recode_WHO_2008 %in% c(7,8,9,10,11,12,13,14,15,16) ~ 2,
                                 AYA_site_recode_WHO_2008 %in% c(17,18,19,20) ~ 3,
                                 AYA_site_recode_WHO_2008 %in% c(21,22,23,24,25) ~ 4,
                                 AYA_site_recode_WHO_2008 %in% c(26,27,28) ~ 5,
                                 AYA_site_recode_WHO_2008 %in% c(29,30) ~ 6,
                                 AYA_site_recode_WHO_2008 %in% c(31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48) ~ 7,
                                 AYA_site_recode_WHO_2008 %in% c(49,50,51,52,53,54,55) ~ 8,
                                 AYA_site_recode_WHO_2008 %in% c(55,56) ~ 9,
                                 AYA_site_recode_WHO_2008 == 99 ~ 10))
                                 
seer0to39$can_typeAYA <- factor(seer0to39$can_typeAYA, levels = c(0:10), labels = c(
  "1. Leukemias",
  "2. Lymphomas",
  "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)",
  "4. Osseous & Chondromatous Neoplasms",
  "5. Soft Tissue Sarcomas",
  "6. Germ Cell and Trophoblastic Neoplasms",
  "7. Melanoma and Skin Carcinomas",
  "8. Carcinomas",
  "9. Miscellaneous specified neoplasms, NOS",
  "10. Unspecified Malignant Neoplasms",
  "Unclassified"))
table(seer0to39$can_typeAYA, useNA = "always")
```
# Exclusions
```{r}
# limit to first cancers
seer0to39 <- seer0to39 %>%
  filter(Firstmalignantprimary_indicator == 1)
```

# Exclude those with missing data on variables used in the analysis
```{r}
seer0to39<-seer0to39 %>%
  filter(!is.na(survmo), !is.na(can_dead), !is.na(In_Medicaid_Flag), !is.na(race_cat), !is.na(age_cat), !is.na(Census_Tract_Poverty_Indicator))


# exclude multiple state Medicaid ids within a single year defined above
seer0to39 <- seer0to39 %>% 
  filter(!patient_id %in% x)
```

# Merge in Derek's variables for number of months pre and post indicator within 3, 9, and 12 month windows
```{r}
# select variables needed from mcseer
mcseer2 <- mcseer %>%
  select(patient_id, status, status3, pre_mo3, post_mo3, status6, pre_mo6, post_mo6, status12, pre_mo12, post_mo12, dxdate, dx_mo_num)

seer0to39<-left_join(seer0to39, mcseer2, by = c("patient_id"))
```


# Create variable for Medicaid continuity with NAs pre or post its own category
```{r}
seer0to39 <- seer0to39 %>%
  mutate(mcont2 = case_when(In_Medicaid_Flag == 0 ~ 0, # Not Linked
                            pre_mo3 >=1 & post_mo3 >= 1 ~ 1, # Linked continuous
                            pre_mo3 > 0 & post_mo3 == 0 ~ 2, # Linked discontinuous Medicaid enrollment 3 mo before or after dx
                            pre_mo3 == 0 & post_mo3 > 0 ~ 3, # Linked discontinuous Medicaid enrollment 3 mo before or after dx
                            pre_mo3 == 0 & post_mo3 == 0 ~ 4, # Linked Not enrolled in Medicaid 3 mo before or after dx
                            is.na(pre_mo3) | is.na(post_mo3)~  5)) # Linked Missing data in 3 mo before and/or after dx
                                        
test3 <- seer0to39 %>%
  group_by(In_Medicaid_Flag, insurance, status3, pre_mo3, post_mo3) %>%
  count(mcont2)

seer0to39$mcont2 <- factor(seer0to39$mcont2, levels = c(0:5), labels = c("Not linked", 
                                                                         "Linked continuous", 
                                                                         "Linked discontinuous enrolled before", 
                                                                         "Linked discontinuous enrolled after",
                                                                         "Linked not enrolled within 3 mo before or 3 mo after dx",
                                                                         "Linked missing data in 3 mo before or after dx"))

table(seer0to39$mcont2, useNA = "always")
table(pre_mo3 =mcseer2$pre_mo3, post_mo3 = mcseer2$post_mo3, useNA = "always")

# Combine categories 2-4 to be discontinuos
seer0to39 <- seer0to39 %>%
  mutate(mcont3 = case_when(mcont2 == "Not linked" ~ 0,
                            mcont2 == "Linked continuous"  ~ 1,
                            mcont2 %in% c("Linked discontinuous enrolled before", "Linked not enrolled within 3 mo before or 3 mo after dx", "Linked not enrolled within 3 mo before or 3 mo after dx", "Linked discontinuous enrolled after") ~2,
                            mcont2 == "Linked missing data in 3 mo before or after dx" ~ 3))


seer0to39$mcont3 <- factor(seer0to39$mcont3, levels = c(0:3), labels = c("Not linked", 
                                                                         "Linked continuous", 
                                                                         "Linked discontinuous ", 
                                                                        "Linked missing data in 3 mo before or after dx"))
table(seer0to39$mcont3, useNA="always")

test <- seer0to39 %>%
  select(patient_id, In_Medicaid_Flag, mcont3, mcont2, pre_mo3, post_mo3) %>%
  filter(In_Medicaid_Flag == "Linked to Medicaid")
```

# make separate dataframes for 0 to 19 and 20 to 39
```{r}
# make separate dataframes for 0 to 19 and 15 to 39
seer0to19 <- seer0to39 %>%
  filter(age_cat %in% c("0-4", "5-9", "10-14", "15-19"))

seer20to39 <- seer0to39 %>%
  filter(age_cat %in% c("20-39"))
```

# Table 1
```{r}
# 0 to 19
library(table1)
label(seer0to39$age_cat)<-"Age category"
label(seer0to39$sex)<-"Sex"
label(seer0to39$insurance)<-"Insurance at diagnosis or first treatment"
label(seer0to39$race_cat)<-"Race category"
label(seer0to39$survmo)<-"Survival months"
label(seer0to39$can_type)<-"Cancer type"

# make dead a factor variable just for this table
seer0to39$can_dead2<-factor(seer0to39$can_dead, levels=c(0,1), labels= c("Alive", "Deceased"))
table(seer0to39$can_dead2)
label(seer0to39$can_dead2) <- "Vital Status"
seer0to39$In_Medicaid_Flag<-factor(seer0to39$In_Medicaid_Flag, levels=c(0,1), labels= c("Not linked to Medicaid", "Linked to Medicaid"))
table(seer0to39$can_dead2)
table(seer0to39$In_Medicaid_Flag)

table1(~age_cat + sex + insurance2 + race_cat + can_dead2 + survmo + can_type|In_Medicaid_Flag, droplevels = TRUE, seer0to39[which(seer0to39$age_cat %in% c("0-4", "5-9", "10-14", "15-19")),])

table1(~age_cat + sex + insurance + race_cat + can_dead2 + survmo + can_typeAYA|In_Medicaid_Flag, droplevels = TRUE, seer0to39[which(seer0to39$age_cat %in% c("20-39")),])
```


# Survival plots (# In_Medicaid_flag)
```{r}
# In_Medicaid_flag
table(seer0to19$insurance)

survp <- function(age_group){
  x <<- seer0to39 %>%
    filter(age_cat %in% age_group)
survival<- ggsurvplot(fit = survfit(Surv(survmo, can_dead) ~ In_Medicaid_Flag, data =x), 
    xlab = "Months from diagnosis", 
    ylab = "Cancer survival proportion",
    legend.title = "",
    legend = "bottom",
    font.legend=c(24, "bold"),
    legend.labs = c("Medicaid not linked", "Medicaid linked"),
    palette=c("black", "blue", "gray", "red"),
    font.x = c(24, "bold"),
    font.y = c(24, "bold"),
    xlim=c(0, 60),
    ylim=c(0.5, 1),
    censor.size=0.5,
    font.tickslab = c(24, "bold"),
    pval=TRUE,
    pval.size = 12,
    pval.coord = c(2,0.55)
    )
}

p <- survp(c("0-4", "5-9", "10-14", "15-19"))
p
p$plot + scale_x_continuous(breaks = c(0,12,24,36,48,60))

ggsave(filename = "KMcurveplot0to19.png",
       width = 14, height = 8.5, units = "in", bg="white") 

p1 <- survp(c("20-39"))
p1
p1$plot + scale_x_continuous(breaks = c(0,12,24,36,48,60))

ggsave(filename = "KMcurveplot20to39.png",
       width = 11, height = 8.5, units = "in", bg="white") 

# estimate survival time 0 to 19
x<-summary(survfit(Surv(survmo, can_dead) ~ In_Medicaid_Flag, data = seer0to19), times = 60)
x

# estimate survival time 20 to 39
y<-summary(survfit(Surv(survmo, can_dead) ~ In_Medicaid_Flag, data = seer20to39), times = 60)
y
```

# Survival plots by age category
```{r}
survp_age <- function(age_group, figtitle){
  x <<- seer0to39 %>%
    filter(age_cat %in% age_group)
survival<- ggsurvplot(fit = survfit(Surv(survmo, can_dead) ~ In_Medicaid_Flag, data =x), 
    xlab = "Months from diagnosis", 
    ylab = "Cancer survival proportion",
    title=figtitle,
    legend.title = "",
    legend = "bottom",
    font.legend=c(12, "bold"),
    legend.labs = c("Medicaid not linked", "Medicaid linked"),
    palette=c("black", "blue", "gray", "red"),
    font.x = c(12, "bold"),
    font.y = c(12, "bold"),
    xlim=c(0, 60),
    ylim=c(0.5, 1),
    censor.size=0.5,
    font.tickslab = c(12, "bold"),
    pval=TRUE,
    pval.size = 4,
    pval.coord = c(2,0.55),
    break.x.by = 12
    )
}
p0to4 <- survp_age("0-4","0-4 years")

p5to9 <- survp_age("5-9", "5-9 years")

p10to14 <- survp_age("10-14", "10-14 years")

p15to19 <- survp_age("15-19", "15-19 years")

library(patchwork)
pdf("KMcurveplot_agecat.pdf", width=11, height=10, paper="USr")
p0to4$plot + p5to9$plot + p10to14$plot + p15to19$plot + plot_layout(ncol=2, nrow=2, guides='collect') &
  theme(legend.position='bottom')
dev.off()

```

# Survival plots by race
```{r}
survp_race <- function(race_f){
  x <<- seer0to19 %>%
    filter(race_ethnicity %in% race_f)
survival<- ggsurvplot(fit = survfit(Surv(survmo, can_dead) ~ In_Medicaid_Flag, data =x), 
    xlab = "Months from diagnosis", 
    ylab = "Cancer survival proportion",
    title=race_f,
    legend.title = "",
    legend = "bottom",
    font.legend=c(12, "bold"),
    legend.labs = c("Medicaid not linked", "Medicaid linked"),
    palette=c("black", "blue", "gray", "red"),
    font.x = c(12, "bold"),
    font.y = c(12, "bold"),
    xlim=c(0, 60),
    ylim=c(0.5, 1),
    censor.size=0.5,
    font.tickslab = c(12, "bold"),
    pval=TRUE,
    pval.size = 4,
    pval.coord = c(2,0.55),
    break.x.by = 12
    )
}
pW <- survp_race("Non-Hispanic White")

pB <- survp_race("Non-Hispanic Black")

pAIAN <- survp_race("Non-Hispanic AIAN")

pAPI <- survp_race("Non-Hispanic API")

pH <- survp_race("Hispanic")

pUn <- survp_race("Non-Hispanic Unknown")

library(patchwork)
pdf("KMcurveplot_race.pdf", width=11, height=10, paper="USr")
pW$plot + pB$plot + pAIAN$plot + pAPI$plot + pH$plot + pUn$plot + plot_layout(ncol=3, nrow=2, guides='collect') &
  theme(legend.position='bottom')
dev.off()

```

# Survival plots by cancer type (Private and Medicaid)
```{r}
# make function for KM curves for age groups and tumor type
KMplots<-function(tumortype, figtitle){
  x<-seer0to19%>%
    filter(can_type == tumortype, insurance %in% c("Private Insurance", "Ever linked to Medicaid"))
  survob <- survfit(
    Surv(survmo, can_dead) ~ In_Medicaid_Flag, data = x)
ggsurvplot(survob, data=x, xlim = c(0,60),  
           pval=TRUE, ylim=c(50,100), fun="pct", 
           size=0.3, 
           conf.int=FALSE, 
           censor=TRUE,  
           title=figtitle,  
           size = 1, 
           censor.size=0.1,
           xlab="Survival time (months)", 
           linetype = 1,
           legend.title="",
           legend.labs = c("Private Insurance", "Medicaid"), palette = c("black", "blue"), break.x.by = 12) +
theme_survminer(font.main = c(20, "bold", "black"), legend=c("bottom"), font.legend=c(20, "bold", "black"), font.y=c(15, "bold"), font.x=c(15, "bold")) 
}

A<-KMplots("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases",  "A. Leukemias, myeloproliferative  diseases, \n and myelodysplastic diseases")
A
B<-KMplots("II. Lymphomas and reticuloendothelial neoplasms",  "b. II. Lymphomas and reticuloendothelial neoplasms")
B
C<-KMplots("III. CNS and miscellaneous intracranial and intraspinal neoplasms",  "B. CNS and miscellaneous intracranial \n and intraspinal neoplasms")
C
D<-KMplots("IV. Neuroblastoma and other peripheral nervous cell tumors",  "d. IV. Neuroblastoma and other peripheral nervous cell tumors")
D
E <- KMplots("V. Retinoblastoma",  "e. V. Retinoblastoma")
E
F <- KMplots("VI. Renal tumors",  "f. VI. Renal tumors")
F
G <- KMplots("VII. Hepatic tumors",  "g. VII. Hepatic tumors")
G
H <- KMplots("VIII. Malignant bone tumors",  "h. VIII. Malignant bone tumors")
H
I <- KMplots("IX. Soft tissue and other extraosseous sarcomas",  "i. IX. Soft tissue and other extraosseous sarcomas")
I
J <- KMplots("X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads",  "j. X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads")
J
K <- KMplots("XI. Other malignant epithelial neoplasms and malignant melanomas",  "k. XI. Other malignant epithelial neoplasms and malignant melanomas")
L <- KMplots("XII. Other and unspecified malignant neoplasms",  "l. XII. Other and unspecified malignant neoplasms")
L

library(patchwork)
pdf("Figure.pdf", width=11, height=10, paper="USr")
A$plot + B$plot + C$plot+ D$plot+ E$plot +F$plot +G$plot +H$plot +I$plot+ J$plot + K$plot + L$plot + plot_layout(ncol=4, nrow=3, guides='collect') &
  theme(legend.position='bottom')
dev.off()

# just leukemia and CNS tumors
pdf("Figure_leukemia_CNS.pdf", width=8.5, height=11, paper="USr")
A$plot + C$plot+ plot_layout(ncol=1, nrow=2, guides='collect') &  theme(legend.position='bottom')
dev.off()

```

# Mcont3 def
```{r}
p<-ggsurvplot(fit = survfit(Surv(survmo, can_dead) ~ mcont3, data =seer0to19), 
    xlab = "Months from diagnosis", 
    ylab = "Cancer survival proportion\n",
    legend.title = "",
    legend.labs = c("Not linked", 
                    "Linked continuous", 
                    "Linked discontinuous ", 
                    "Linked missing data in 3 \n mo before or after dx"),
    font.legend=c(20, "bold"),
    legend = c("bottom"),
    palette=c("black", "maroon2", "cyan3", "red", "green", "blue"),
    font.x = c(24, "bold"),
    font.y = c(24, "bold"),
    xlim=c(0, 60),
    ylim=c(0.5, 1),
    censor.size=0.5,
    break.x.by = 12,
    font.tickslab = c(24, "bold"),
    pval=TRUE,
    pval.size = 12,
    pval.coord = c(2,0.55),
    font.title = c(34, "bold"),
    risk.table = FALSE,
    risk.table.fontsize=3,
    tables.height = 0.2,
    tables.theme = theme_cleantable())
p

p$plot + scale_x_continuous(breaks = c(0,12,24,36,48,60))
ggsave(filename = "Continuity 3 month KM plot 0 to 19.png",
       width = 14, height = 8.5, units = "in", bg="white") 


p1<-ggsurvplot(fit = survfit(Surv(survmo, can_dead) ~ mcont3, data =seer20to39), 
    xlab = "Months from diagnosis", 
    ylab = "Cancer survival proportion",
    legend.title = "",
    #legend.labs = c("Not linked", 
                    #"Linked continuous", 
                    #"Linked discontinuous enrolled before", 
                    #"Linked discontinuous enrolled after",
                    #"Linked not enrolled within 3 mo before or 3 mo after dx",
                    #"Linked missing data in 3 mo before or after dx"),
    font.legend=c(8, "bold"),
    legend = c("right"),
    palette=c("black", "maroon2", "cyan3", "red", "green", "blue"),
    font.x = c(18, "bold"),
    font.y = c(18, "bold"),
    xlim=c(0, 60),
    ylim=c(0, 1),
    censor.size=0.5,
    font.tickslab = c(16, "bold"),
    pval=TRUE,
    risk.table = TRUE,
    risk.table.fontsize=3,
    tables.height = 0.2,
    tables.theme = theme_cleantable())
p1

ggsave(filename = "Continuity 3 month KM plot 20 to 39.png",
       width = 11, height = 8.5, units = "in", bg="white") 

```

# make function for KM curves for age groups and tumor type
```{r}
KMplots<-function(tumortype, age_group, figtitle){
  x<<-seer0to39 %>%
    filter(can_type == tumortype, age_cat == age_group)
  survob <- survfit(
    Surv(survmo, can_dead) ~ mcont3, data = x)
ggsurvplot(survob, data=x, xlim = c(0,60),  
           pval=TRUE, ylim=c(0,100), fun="pct", 
           size=0.3, 
           conf.int=FALSE, 
           censor=TRUE,  
           title=figtitle,  
           size = 1, 
           censor.size=0.1,
           xlab="Survival time (months)", 
           linetype = 1,
           legend.title="",
           #legend.labs = c("Private Insurance", "Medicaid"), 
           palette = c("black", "maroon2", "cyan3", "red", "green", "blue"), break.x.by = 12) +
theme_survminer(font.main = c(20, "bold", "black"), legend=c("right"), font.legend=c("bold", "black"), font.y=c(15, "bold"), font.x=c(15, "bold")) 
}

A<-KMplots("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", c("0-4", "5-9", "10-14", "15-19"), "A. Leukemias, myeloproliferative  diseases, \n and myelodysplastic diseases")
A

B<-KMplots("II. Lymphomas and reticuloendothelial neoplasms", c("0-4", "5-9", "10-14", "15-19"), "b. II. Lymphomas and reticuloendothelial neoplasms")
B

C<-KMplots("III. CNS and miscellaneous intracranial and intraspinal neoplasms", c("0-4", "5-9", "10-14", "15-19"),  "B. CNS and miscellaneous intracranial \n and intraspinal neoplasms")
C

D<-KMplots("IV. Neuroblastoma and other peripheral nervous cell tumors", c("0-4", "5-9", "10-14", "15-19"), "d. IV. Neuroblastoma and other peripheral nervous cell tumors")
D
#E <- KMplots("V. Retinoblastoma",  "e. V. Retinoblastoma")
#E

F <- KMplots("VI. Renal tumors", c("0-4", "5-9", "10-14", "15-19"), "f. VI. Renal tumors")
F

G <- KMplots("VII. Hepatic tumors", c("0-4", "5-9", "10-14", "15-19"), "g. VII. Hepatic tumors")
G

H <- KMplots("VIII. Malignant bone tumors", c("0-4", "5-9", "10-14", "15-19"), "h. VIII. Malignant bone tumors")
H

I <- KMplots("IX. Soft tissue and other extraosseous sarcomas", c("0-4", "5-9", "10-14", "15-19"), "i. IX. Soft tissue and other extraosseous sarcomas")
I

J <- KMplots("X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads", c("0-4", "5-9", "10-14", "15-19"), "j. X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads")
J

K <- KMplots("XI. Other malignant epithelial neoplasms and malignant melanomas", c("0-4", "5-9", "10-14", "15-19"), "k. XI. Other malignant epithelial neoplasms and malignant melanomas")
K

L <- KMplots("XII. Other and unspecified malignant neoplasms", c("0-4", "5-9", "10-14", "15-19"), "l. XII. Other and unspecified malignant neoplasms")
L

library(patchwork)
pdf("Figure.pdf", width=11, height=10, paper="USr")
A$plot + B$plot + C$plot+ D$plot+F$plot +G$plot +H$plot +I$plot+ J$plot + K$plot + L$plot + plot_layout(ncol=4, nrow=3, guides='collect') &
  theme(legend.position='bottom')
dev.off()

table(seer0to19$can_type, seer0to19$mcont2)

```
# Cox proportional hazards models (0 to 19) San Antonio conference
```{r}
# unadjusted model 0 TO 19
res.cox1 <- coxph(Surv(survmo, can_dead) ~ In_Medicaid_Flag, data = seer0to19) # Univariable model
res.cox1
summary(res.cox1)

# adjusted models 0 to 19
res.cox2 <- coxph(Surv(survmo, can_dead) ~ In_Medicaid_Flag +race_cat +age_cat + Census_Tract_Poverty_Indicator, data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox2
summary(res.cox2)

# mcont3 unadjusted model 0 TO 19
res.cox3 <- coxph(Surv(survmo, can_dead) ~ mcont3, data = seer0to19) # Univariable model
res.cox3
summary(res.cox3)

# mcont3 adjusted models 0 to 19
res.cox4 <- coxph(Surv(survmo, can_dead) ~ mcont3 +race_cat +age_cat + Census_Tract_Poverty_Indicator, data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox4
summary(res.cox4)

a<-as.data.frame(exp(cbind(HR = coef(res.cox1), confint(res.cox1))))
   a<-a[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="1. Unadjusted") %>%
    mutate(age_group = "0 to 19 years") 
   
b<-as.data.frame(exp(cbind(HR = coef(res.cox2), confint(res.cox2))))
   b<-b[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="2. Adjusted for race, age, cenus tract poverty indicator") %>%
    mutate(age_group = "0 to 19 years")

c<-as.data.frame(exp(cbind(HR = coef(res.cox3), confint(res.cox3))))
   c<-c[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model="3. Unadjusted") %>%
    mutate(age_group = "0 to 19 years")

d<-as.data.frame(exp(cbind(HR = coef(res.cox4), confint(res.cox4))))
   d<-d[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model="4. Adjusted for race, age, cenus tract poverty indicator") %>%
    mutate(age_group = "0 to 19 years")
SA_results <- rbind(a,b,c,d)  
```

# Cox proportional hazards models (20 to 39) ASCO
```{r}
# unadjusted model 20 to 39
res.cox1 <- coxph(Surv(survmo, can_dead) ~ In_Medicaid_Flag, data = seer20to39) # Univariable model
res.cox1
summary(res.cox1)

# adjusted models 20 to 39
res.cox2 <- coxph(Surv(survmo, can_dead) ~ In_Medicaid_Flag +race_cat + Census_Tract_Poverty_Indicator, data = seer20to39) # Model adjusted for race age and a measure of poverty
res.cox2
summary(res.cox2)

# mcont3 unadjusted model 20 to 39
res.cox3 <- coxph(Surv(survmo, can_dead) ~ mcont3, data = seer20to39) # Univariable model
res.cox3
summary(res.cox3)

# mcont3 adjusted models 20 to 39
res.cox4 <- coxph(Surv(survmo, can_dead) ~ mcont3 +race_cat  + Census_Tract_Poverty_Indicator, data = seer20to39) # Model adjusted for race age and a measure of poverty
res.cox4
summary(res.cox4)

a<-as.data.frame(exp(cbind(HR = coef(res.cox1), confint(res.cox1))))
   a<-a[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="1. Unadjusted") %>%
    mutate(age_group = "20 to 39 years") 
   
b<-as.data.frame(exp(cbind(HR = coef(res.cox2), confint(res.cox2))))
   b<-b[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="2. Adjusted for race, age, cenus tract poverty indicator") %>%
    mutate(age_group = "20 to 39 years")

c<-as.data.frame(exp(cbind(HR = coef(res.cox3), confint(res.cox3))))
   c<-c[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model="3. Unadjusted") %>%
    mutate(age_group = "20 to 39 years")

d<-as.data.frame(exp(cbind(HR = coef(res.cox4), confint(res.cox4))))
   d<-d[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model="4. Adjusted for race, age, cenus tract poverty indicator") %>%
    mutate(age_group = "20 to 39 years")

ASCO <- rbind(a,b,c,d)  
```
# Examine effect modification by cqncer type
```{r}

res.cox2 <- coxph(Surv(survmo, can_dead) ~ In_Medicaid_Flag +race_cat + Census_Tract_Poverty_Indicator + can_typeAYA, data = seer20to39) # Model adjusted for race age and a measure of poverty
res.cox2
summary(res.cox2)


res.cox2em <- coxph(Surv(survmo, can_dead) ~ In_Medicaid_Flag +race_cat + Census_Tract_Poverty_Indicator + can_typeAYA + can_typeAYA*In_Medicaid_Flag, data = seer20to39) # Model adjusted for race age and a measure of poverty
res.cox2
summary(res.cox2em)

library(lmtest)
lrtest(res.cox2, res.cox2em)

table(seer20to39$can_typeAYA)
table(seer20to39$age_cat)

```

# get results for each cancer type for In_Medicaid_Flag
```{r}
cox <-function(tumortype){
  x<<-seer20to39 %>%
    filter(can_typeAYA == tumortype)
    
coxph(Surv(survmo, can_dead) ~ In_Medicaid_Flag + race_cat+ Census_Tract_Poverty_Indicator, data = x) # Model adjusted for race  and a measure of poverty
}

a <- summary(cox("1. Leukemias"))
a

b <- summary(cox("2. Lymphomas"))
b

c <- summary(cox("3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)"))
c

d <- summary(cox("4. Osseous & Chondromatous Neoplasms"))
d

e <- summary(cox("5. Soft Tissue Sarcomas"))
e

f <- summary(cox("6. Germ Cell and Trophoblastic Neoplasms"))
f

g <- summary(cox("7. Melanoma and Skin Carcinomas"))
g

h <- summary(cox("8. Carcinomas"))
h

i <- summary(cox("9. Miscellaneous specified neoplasms, NOS"))
i
```
# get results for each cancer type for mcont3
```{r}
cox <-function(tumortype){
  x<<-seer20to39 %>%
    filter(can_typeAYA == tumortype)
    
coxph(Surv(survmo, can_dead) ~ mcont3 + race_cat+ Census_Tract_Poverty_Indicator, data = x) # Model adjusted for race  and a measure of poverty
}

a <- summary(cox("1. Leukemias"))
a

b <- summary(cox("2. Lymphomas"))
b

c <- summary(cox("3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)"))
c

d <- summary(cox("4. Osseous & Chondromatous Neoplasms"))
d

e <- summary(cox("5. Soft Tissue Sarcomas"))
e

f <- summary(cox("6. Germ Cell and Trophoblastic Neoplasms"))
f

g <- summary(cox("7. Melanoma and Skin Carcinomas"))
g

h <- summary(cox("8. Carcinomas"))
h

i <- summary(cox("9. Miscellaneous specified neoplasms, NOS"))
i
```

# Export tables to excel
```{r}

# library(xlsx)
write.xlsx(SA_results, file="analysis_tables.xlsx", sheetName = "Table Cox PH 0 to 19", showNA=TRUE, append = TRUE)
write.xlsx(ASCO, file="analysis_tables.xlsx", sheetName = "Table Cox PH 20 to 39", showNA=TRUE, append = TRUE)
```

# 20-39 ASCO abstract
```{r}
# adjusted models 20 to 39
res.cox <- coxph(Surv(survmo, can_dead) ~ In_Medicaid_Flag + Age_recode_with_1_year_olds + race_cat + Census_Tract_Poverty_Indicator, data = seer20to39) # Model adjusted for race age and a measure of poverty
res.cox
summary(res.cox)

# adjusted models 20 to 39
res.cox <- coxph(Surv(survmo, can_dead) ~ mcont3 +race_cat + Age_recode_with_1_year_olds + Census_Tract_Poverty_Indicator, data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox
summary(res.cox)

# look at cancer types 
leuk <-seer20to39 %>%
  filter(can_type == "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases")

res.cox <- coxph(Surv(survmo, can_dead) ~ mcont3 +race_cat + Age_recode_with_1_year_olds + Census_Tract_Poverty_Indicator, data = leuk) # Model adjusted for race age and a measure of poverty
res.cox
summary(res.cox)
```

```{r}
# adjusted model for leukemia and CNS tumors
leuk <-seer0to19 %>%
  filter(can_type == "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", insurance %in% c("Private Insurance", "Medicaid"))
  
leuk.cox <- coxph(Surv(survmo, can_dead) ~ insurance +race_cat +age_cat + Census_Tract_Poverty_Indicator, data = leuk) # Model adjusted for race age and a measure of poverty
leuk.cox
summary(leuk.cox)

brain <-seer0to19 %>%
  filter(can_type == "III. CNS and miscellaneous intracranial and intraspinal neoplasms", insurance %in% c("Private Insurance", "Medicaid"))

brain.cox <- coxph(Surv(survmo, can_dead) ~ insurance +race_cat +age_cat + Census_Tract_Poverty_Indicator, data = brain) # Model adjusted for race age and a measure of poverty
brain.cox
summary(brain.cox)

# Interactions
# race_cat
res.cox2 <- coxph(Surv(survmo, can_dead) ~ In_Medicaid_Flag +race_cat +age_cat + Census_Tract_Poverty_Indicator + race_cat*In_Medicaid_Flag, data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox2
summary(res.cox2)

# census tract poverty indicator
res.cox3 <- coxph(Surv(survmo, can_dead) ~ In_Medicaid_Flag +race_cat +age_cat + Census_Tract_Poverty_Indicator + Census_Tract_Poverty_Indicator*In_Medicaid_Flag, data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox3
summary(res.cox3)
```

# Overall for San Antonio conference
```{r}
# census tract poverty indicator
res.cox4 <- coxph(Surv(survmo, can_dead) ~ mcont3 +race_cat + Census_Tract_Poverty_Indicator, data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox4
summary(res.cox4)
```

# Conduct analyses using the 3 month variable
```{r}
res.cox <- coxph(Surv(survmo, can_dead) ~ enroll_cat +race_cat +age_cat + Census_Tract_Poverty_Indicator, data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox
summary(res.cox)

# what is time to event in those enrolled Not enrolled in 3 month window before but not after?

three_before <- seer0to19 %>%
  group_by(enroll_cat, can_dead) %>%
  summarise(med = median(survmo), mean = mean(survmo), n = n())

table(three_before$can_dead)
```

